Sulfur Dioxide on Water's Surface

### 2011-04-04 TransitOrientation.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. There are differences between where the orientation transitions from isotropic in the neat-water and saturated systems. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.

  % talk about theta differences
  The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase. On approach to the neat-water surface, that transition occurs once just above 10\angs~from the surface, and again at 5\angs~from the surface and below. The onset of a preferred orientation at 10\angs~occurrs at the cutoff length of the simulated intermolecular bonding interactions. Between 5-10\angs~the molecule appears to orient more isotropically. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx. 8\angs~above the surface location. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few\angs~within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.

  % talk about phi differences
  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis. This is the case in both systems with one exception in the neat-water system. At the water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$. This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal). However, below this 5\angs~region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

@@

	%Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface. While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system. The orientation effect is shallower under the neat-\wat~surface. However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one. The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system. In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.

>>>>

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. There are differences between the two systems as to where the point of transition from isotropic to preferred orientations happens. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.

  % talk about theta differences
  The $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic. There is a transition below 10\angs~where the \suldiox~orients with its bisector perpendicular to the water surface. On approach to the neat-water surface, that transition occurs at approximately 5\angs~from the surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few\angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
		\caption{(A) \suldiox~molecules above and below both neat-water and saturated surfaces orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		\label{fig:so2-surface-angles}
	\end{center}
\end{figure}

  % talk about phi differences
  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis. This is the case in both systems, with one exception in the neat-water system. At the neat-water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$. This corresponds to a \suldiox~orienting less flat to the surface (inclined up to 60 degrees from the surface normal). The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinate bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

@@

	%Differences ocurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface. While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system. The orientation effect is shallower under the neat-\wat~surface. However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one. The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system. In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.
